Hierarchical Inference of the Intrinsic Ellipticity Distribution of Galaxies during Weak Lensing Analysis

### 2013-04-07 docs/ms.tex
Update ms.tex

Weak gravitational lensing has the potential to be the most powerful
tool to explore the dark universe. One application of weak gravitational lensing is
that it can be used to measure the mass profile of single galaxy,galaxy
groups , clusters  the large scales. Cosmic shear is the mild distortion
of distant galaxy images due to the bending of light by intervening
matter.It provides one of the most promising methods for constraining
the nature of dark energy (Albrecht et al. 2006; Peacock\&Schneider
2006). While in galaxy--galaxy weaklensing, tangential shear distortions
of background galaxies around foreground galaxies allows direct measurement
of the galaxy-DM correlation.

@@

A typical procedure of weak lensing analysis  includes galaxy shape measurements from images
,shear inference and mass profile reconstruction. The galaxy shape measurement is usually the
first step of weak lensing analysis,since it's the direct observable related with shear effect. However
in this paper we only concern about how to infer shear from galaxy
ellipticity in a novel hierarchical way. Therefore we set up a simple toy model in which we have the observed ellipticity as
data,and the only shape measurement error we consider is a Gaussian noise
added on the lensed galaxy ellipticity. We also suppose there is a uniform shear
acting on one all the galaxies in one sky patch.

@@

This paper is orgnized as following: In section 2 we introduce the
weak lensing formulas and the likelihood functions. Then we describe
two flexible models for the intrinsic shape distribution.Section 3 details
the results of simulations including tests for biases and accuracy.In Section 4, we
derive some general conclusions and outline suggestions for future
development.

@@

If there is measurement error on $\epsilon_{ob}$,we will have 
\begin{equation}
p(\epsilon_{ob}|\vec{g},\alpha)=\int f(\epsilon_{ob}|\epsilon_{\ell})\, p(\epsilon_{\ell}|\vec{g},\alpha)\,d\epsilon_{\ell}
\end{equation}
Here the prior distribtuion we use for $\vec{g}$ is a 2D flat distribution
in the region $|\vec{g}|<1$.Therefore $p(\vec{g)}=1$ if $|\vec{g}|<1$
and $p(\vec{g})=0$ if $|\vec{g}|>1$.

@@

Since $p(\alpha)$ does not explicitly depends on $\vec{g}$ .$p(\alpha|\vec{g})$is
just the pror distribution $p(\vec{\alpha}|\xi)$.Therefore, we have
marginalized likehood 
\begin{equation}
p(\epsilon_{\ell}|\vec{g})=\int p(\epsilon_{\ell}|\vec{g},\vec{\alpha})\, p(\vec{\alpha}|\xi)\,d\vec{\alpha}
\end{equation}
Given that marginalized likelihood function, in principle we could
optimize it and gives out a max-likelihood estimator for $\vec{g}$
in different patches simultaneously.

One thing to note is that the likelihood of g in patch i only depends
on the shape parameters but not dependent to g in other patch.Therefore,the
marginalized likelihood of shape parameters has an equation like:

@@

Here $\xi$ is a prefactor need to be tuned .

Meanwhile,in order to show the correctness of our method, we also
set up a simulation in which we know the true $P(|\epsilon_{0}||\alpha)$
is a Beta distribution, and thus $\alpha$ is the 2-D parameter (p,q)
of the Beta distribution.We will show the posterior samples using
both of these two methods in next section. Note that for either model
we assume P($\epsilon_{0}$) is isotropic.

@@

One thing to note is that the conditional probability of g in ith
patch only explicitly depends on $\alpha$, but not g in other patches. Therefore
with $\alpha$ fixed ,the posterior distribution of reduced shear
in ith patch are independent from each other. 
\begin{equation}
p(\vec{g}_{1}....\vec{g}_{n}|\alpha,Data)=p(\vec{g}_{1}|\alpha,Data_{1})....p(\vec{g}_{n}|\alpha,Data_{n})
\end{equation}
Therefore sampling reduced shear g from conditional distribution can
be reduced to 2D sampling problem,which can be done very quickly.

@@

The procedure of our simulation is that ,first we generated N{*}NP
unlensed galaxies' elliprticity from the ``true'' $P(\epsilon_{0})$,and
asign these galaxies into NP patches equally.So each patch has N galaxies.Second
we generate NP true reduced shears ,and asign them to each patch,and
then shear the unlensed galaxies using formula (4) .

@@

The probability distribution can be visualised by plotting points at each sample
position. An example in the x, y plane is shown for the posterior distribution of
$\vec{g}$ of 64 sky patches in Fig. .8 galaxy shape profiles in each patch is used as the data.
The number density of samples is proportional to the probability density.The figure shows that the two parameters of $\vec{g}$
do not have strong correlation,since the shape of smaples is close to a circle.
We also shows the histogram of g1 in a single sky patch. Not surprisingly it is close to a Gaussian distribution.
If the probability distribution of parameters are perfect Gaussian,their most probable estimates and uncertainty 
is just the mean and standard deviation of the Gaussian distribution.However,in pratical,we find that the distribution 
of reduced shear $\vec{g}$ will be getting closer to Gaussian,with larger number of galaxies used.When we have more than
8 galaxies in each sky patch,we found it's eligible to make such an approximation that to use sample mean and standard 
deviation as most probable estimates and uncertainty. We will justify this approximation with a bias test in the following
sub section.

\subsection{Bias Test}
In this section,we measured the levels of possible systemmatical
bias of this Hierarchical inference method.In principle given the correct model and prior ,
the hierchical Bayesian inference method would have no bias, if propagate the whole likelihood function.
However,in practical, the bias may be introduced by the model we use for the intrinsic shape distribution,
and the prior regularizer of the prefactor $\xi$ as well as the Gaussian approximation we did for the likelihood
function of $\vec{g}$. Therefore a bias test is necessary to identify how much bias there is. 
Following Heymans et al. (2006), we quantify the systematic bias in
in terms of the multiplicative error m and additive error c on the
true shear $g^{true}$ 
\begin{equation}
\hat{g_{i}}-g_{i}^{true}=m_{i}g_{i}^{true}+c_{i}
\end{equation}

@@

We compare the mean square error of our estimator with average and
the estimator that could be constructed if the true unlensed ellipticity
distribution were known. We show how the mean square error varies
with number of galaxies ,number of patches and so on.

\subsection{Measurement Error}
In this section we  discuss how the shape measurement error will
affect the variance and bias of this Hierarchical Bayesian method.
In principle,if the likelihood of true lensed ellipticity $p(\epsilon_{ob}|\epsilon_{\ell})$
is known the likelihood $p(\epsilon_{ob}|g,\alpha)$ can be calculated using the convolution as equation ().
However,due to the limitation of computational time we avoid to do this convolution but use $\epsilon_{ob}$
as an estimator of $\epsilon_{\ell}$.
The likelihood of true lensed ellipticity $p(\epsilon_{ob}|\epsilon_{\ell})$ we use in the simulation is a 2D
Gaussian with mean equals $\epsilon_{\ell}$ and $\sigma$ equals 0.05 .
The result of bias test and mean square error test is shown in...

>>>>

Weak gravitational lensing has the potential to be the most powerful
tool to explore the dark universe. One application of weak gravitational lensing is
that it can be used to measure the mass profile of single galaxy,galaxy
groups , clusters  the large scales. Cosmic shear is the mild distortion
of distant galaxy images due to the bending of light by intervening
matter. It provides one of the most promising methods for constraining
the nature of dark energy (Albrecht et al. 2006; Peacock\&Schneider
2006). While in galaxy--galaxy weaklensing, tangential shear distortions
of background galaxies around foreground galaxies allows direct measurement
of the galaxy-DM correlation.

@@

A typical procedure of weak lensing analysis  includes galaxy shape measurements from images
,shear inference and mass profile reconstruction. The galaxy shape measurement is usually the
first step of weak lensing analysis, since it's the direct observable related with shear effect. However
in this paper we only concern about how to infer shear from galaxy
ellipticity in a novel hierarchical way. Therefore we set up a simple toy model in which we have the observed ellipticity as
data, and the only shape measurement error we consider is a Gaussian noise
added on the lensed galaxy ellipticity. We also suppose there is a uniform shear
acting on one all the galaxies in one sky patch.

@@

This paper is orgnized as following: In section 2 we introduce the
weak lensing formulas and the likelihood functions. Then we describe
two flexible models for the intrinsic shape distribution. Section 3 details
the results of simulations including tests for biases and accuracy. In Section 4, we
derive some general conclusions and outline suggestions for future
development.

@@

If there is measurement error on $\epsilon_{ob}$,we will have 
\begin{equation}
p(\epsilon_{ob}|\vec{g},\alpha)=\int f(\epsilon_{ob}|\epsilon_{\ell})\, p(\epsilon_{\ell}|\vec{g},\alpha)\,d\epsilon_{\ell}
\end{equation}
Here the prior distribtuion we use for $\vec{g}$ is a 2D flat distribution
in the region $|\vec{g}|<1$. Therefore $p(\vec{g)}=1$ if $|\vec{g}|<1$
and $p(\vec{g})=0$ if $|\vec{g}|>1$.

@@

Since $p(\alpha)$ does not explicitly depends on $\vec{g}$ .$p(\alpha|\vec{g})$is
just the pror distribution $p(\vec{\alpha}|\xi)$. Therefore, we have
marginalized likehood 
\begin{equation}
p(\epsilon_{\ell}|\vec{g})=\int p(\epsilon_{\ell}|\vec{g},\vec{\alpha})\, p(\vec{\alpha}|\xi)\,d\vec{\alpha}
\end{equation}
Given that marginalized likelihood function, in principle we could
optimize it and gives out a max-likelihood estimator for $\vec{g}$
in different patches simultaneously.

One thing to note is that the likelihood of g in patch i only depends
on the shape parameters but not dependent to g in other patch. Therefore, the
marginalized likelihood of shape parameters has an equation like:

@@

Here $\xi$ is a prefactor need to be tuned.

Meanwhile,in order to show the correctness of our method, we also
set up a simulation in which we know the true $P(|\epsilon_{0}||\alpha)$
is a Beta distribution, and thus $\alpha$ is the 2-D parameter (p,q)
of the Beta distribution. We will show the posterior samples using
both of these two methods in next section. Note that for either model
we assume P($\epsilon_{0}$) is isotropic.

@@

One thing to note is that the conditional probability of g in ith
patch only explicitly depends on $\alpha$, but not g in other patches. Therefore
with $\alpha$ fixed ,the posterior distribution of reduced shear
in ith patch are independent from each other. 
\begin{equation}
p(\vec{g}_{1}....\vec{g}_{n}|\alpha,Data)=p(\vec{g}_{1}|\alpha,Data_{1})....p(\vec{g}_{n}|\alpha,Data_{n})
\end{equation}
Therefore sampling reduced shear g from conditional distribution can
be reduced to 2D sampling problem, which can be done very quickly.

@@

The procedure of our simulation is that, first we generated N{*}NP
unlensed galaxies' elliprticity from the ``true'' $P(\epsilon_{0})$,and
asign these galaxies into NP patches equally. So each patch has N galaxies.Second
we generate NP true reduced shears, and asign them to each patch,and
then shear the unlensed galaxies using formula (4) .

@@

The probability distribution can be visualised by plotting points at each sample
position. An example in the x, y plane is shown for the posterior distribution of
$\vec{g}$ of 64 sky patches in Fig. .8 galaxy shape profiles in each patch is used as the data.
The number density of samples is proportional to the probability density.The figure shows that the two parameters of $\vec{g}$
do not have strong correlation,since the shape of smaples is close to a circle.
We also shows the histogram of g1 in a single sky patch. Not surprisingly it is close to a Gaussian distribution.
If the probability distribution of parameters are perfect Gaussian, the most probable estimates and uncertainty 
is just the mean and standard deviation of the Gaussian distribution. However,in pratical,we find that the distribution 
of reduced shear $\vec{g}$ will be getting closer to Gaussian,with larger number of galaxies used. When we have more than
8 galaxies in each sky patch, we found it's eligible to make such an approximation that to use sample mean and standard 
deviation as most probable estimates and uncertainty. We will justify this approximation with a bias test in the following
sub section.

\subsection{Bias Test}
In this section,we measured the levels of possible systemmatical
bias of this Hierarchical inference method.In principle given the correct model and prior ,
the hierchical Bayesian inference method would have no bias, if propagate the whole likelihood function.
However, in practical, the bias may be introduced by the model we use for the intrinsic shape distribution,
and the prior regularizer of the prefactor $\xi$ as well as the Gaussian approximation we did for the likelihood
function of $\vec{g}$. Therefore a bias test is necessary to identify how much bias there is. 
Following Heymans et al. (2006), we quantify the systematic bias in
in terms of the multiplicative error m and additive error c on the
true shear $g^{true}$ 
\begin{equation}
\hat{g_{i}}-g_{i}^{true}=m_{i}g_{i}^{true}+c_{i}
\end{equation}

@@

We compare the mean square error of our estimator with average and
the estimator that could be constructed if the true unlensed ellipticity
distribution were known. We show how the mean square error varies
with number of galaxies, number of patches and so on.

\subsection{Measurement Error}
In this section we  discuss how the shape measurement error will
affect the variance and bias of this Hierarchical Bayesian method.
In principle, if the likelihood of true lensed ellipticity $p(\epsilon_{ob}|\epsilon_{\ell})$
is known the likelihood $p(\epsilon_{ob}|g,\alpha)$ can be calculated using the convolution as equation ().
However,due to the limitation of computational time we avoid to do this convolution but use $\epsilon_{ob}$
as an estimator of $\epsilon_{\ell}$.
The likelihood of true lensed ellipticity $p(\epsilon_{ob}|\epsilon_{\ell})$ we use in the simulation is a 2D
Gaussian with mean equals $\epsilon_{\ell}$ and $\sigma$ equals 0.05 .
The result of bias test and mean square error test is shown in...
